Concurrent Programming with CXL

### 2020-05-15 book/paper.tex
Latest

\title{Introduction Concurrent Programming with CXL}
\author{RVR and others}

@@

\chapter{Introduction to CXL}

@@

\chapter{Naive Solutions}

@@

>>>>

\title{Concurrent Programming with CXL}
\author{RVR and others}

@@

\chapter{Introduction to Programming with CXL}

@@

\chapter{Naive Attempts at a Solution}

\begin{figure}
\begin{code}
\verbatiminput{UpLock.cxl}
\end{code}
\caption{Naive implementation of a lock.}
\label{fig:uplock}
\end{figure}

\begin{figure}
\begin{code}
\verbatiminput{UpFlags.cxl}
\end{code}
\caption{Naive use of flags to solve mutual exclusion.}
\label{fig:upflags}
\end{figure}

\begin{figure}
\begin{code}
\begin{verbatim}
==== Livelock ====
__main__ 0 [0,35-63]    64 {'flags': {2: False, 1: False}, 'count': 0,
                                'done': {2: False, 1: False}}
up       1 [1-13,21-24] 25 {'flags': {2: False, 1: True}, 'count': 0,
                                'done': {2: False, 1: False}}
\end{verbatim}
\end{code}
\caption{Error output of Figure~\ref{fig:upflags}.}
\label{fig:upflagsoutput}
\end{figure}

@@

Figure~\ref{fig:uplock} presents a solution based on a \emph{lock}.
The idea is that the lock is a baton that at most one process can own at
a time.
Initially the lock is not owned, indicated by lock being \texttt{False}.
To enter the critical section, a process waits until lock is \texttt{False}
and then sets it to \texttt{True} to indicate that the lock has been taken.
The process then executes the critical section.  Finally the process
releases the lock by setting it back to \texttt{False}.

Unfortunately, if we run the program, we find that the assertion still
fails.  Diagnosing the problem, we see that the \texttt{lock} variable
suffers from the same problem as the \texttt{count} variable: operations
on it consist of several instructions are divisible.  It is thus possible
for both processes to believe the lock is available and obtain the lock
at the same time.

Figure~\ref{fig:upflags} presents a solution based on each process having
a flag indicating that it is trying to enter the critical section.
After setting the flag, the process waits until the other process is not
trying to enter the critical section.
If we run this program, the assertion does not fail.  In fact, this
solution does prevent both processes being in the critical section at
the same time.
However, it has a different problem called livelock: if both try to
enter the critical section at the same time, they may end up waiting for
one another.
